---
title: "Points vs Shooting Percentage"
author: "Will Ives"
date: "February 23, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(plotly)
library(xlsx)
library(XML)
```

## Theory
In the NHL there is a theory that the shooting percentage is fairly consistant for a player in the long run, but highly fluctuates in the short run.  If this is correct then we would expect many of the top goal-scorers to have an unusally high shooting percentage during the season. Therefore we will look at the graph of shooting percentage vs goals during the 2017 season.

``` {r pull in data, echo = FALSE, warnings = FALSE }
URL <- "http://www.hockey-reference.com/leagues/NHL_2017_skaters.html"
    tables <- readHTMLTable(URL, header=FALSE)
    ## for tutorial sake -- its the first table in the list
    cur <- tables[[1]]
    
    ## set columns to characters and set column names
    cur <- data.frame(lapply(cur, as.character),stringsAsFactors = FALSE)
    c<- c("Rk","Player","Age","Pos","Tm","GP","G","A","PTS","P.M","PIM","EVG","PPG","SHG","GWG","EVA","PPA","SHA","S","Sp","TOI","ATOI","BLK","HIT","FOW","FOL","FOp" )
    colnames(cur) <- c
    ## remove the column breaks
    cur <- cur[cur$Rk!='Rk',]
    
    ## holds the column indexes for the loop
    index <- c(1,3,6:(ncol(cur)))
    ## change to numeric
    for(i in index) {
      cur[,i] <- as.numeric(cur[,i])
    }
    
    # Fix ATOI and NA's
    cur$ATOI <- cur$TOI/cur$GP
    cur[is.na(cur)] <-0
    
    # Change Team and Position to factors
    cur$Tm <- as.factor(cur$Tm)
    cur$Pos <- as.factor(cur$Pos)
    cur <- cur[with(cur, order(-G)), ]
    topp <- head(cur,100)
```

## Plot of Shots vs Goals in 2017
```{r plot the graph, warning = FALSE, echo = FALSE, message = FALSE}
p <- plot_ly(topp, x = ~G, y = ~Sp, type = 'scatter', mode = 'markers',
        text = ~Player)
p
```

## Conclusion
As you can see in the plot there appears to be a group of players with shooting percentages above 20 that are likely to not score as many goals in the second half of the season as the first.